### 2013-06-26 paper/supplemental.tex
moved big table to separate file

\documentclass[12pt]{article}

\usepackage[margin=1in]{geometry}  % set the margins to 1in on all sides
\usepackage{deluxetable}             
\usepackage{graphicx}              % to include figures
\usepackage{amsmath}               % great math stuff
\usepackage{amsfonts}              % for blackboard bold, etc
\usepackage{amsthm}                % better theorem environments
\usepackage{url}

% various theorems, numbered by section

\newtheorem{thm}{Theorem}[section]
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{conj}[thm]{Conjecture}
\newtheorem{dfn}[thm]{Definition}

\DeclareMathOperator{\id}{id}

\newcommand{\bd}[1]{\mathbf{#1}}  % for bolding symbols
\newcommand{\RR}{\mathbb{R}}      % for Real numbers
\newcommand{\ZZ}{\mathbb{Z}}      % for Integers
\newcommand{\col}[1]{\left[\begin{matrix} #1 \end{matrix} \right]}
\newcommand{\comb}[2]{\binom{#1^2 + #2^2}{#1+#2}}

\begin{document}

\title{A Generalized Sphere Theorem for Positively Curved Combinatorial 3-Manifolds: Supplemental Information}

\maketitle

\section{Table of Diameters}

The following table lists the diameter of each of the positively curved 3-manifolds.  Here $N$ refers to the manifold number as tabulated by Lutz, Sullivan and Sulankein (see citations in paper), $t$ refers to the topological type, and $d$ refers to diameter, when hops ($H = \frac{2}{3}\sqrt{2}$) and jumps ($J = \frac{1}{2}\sqrt{2} + \sqrt{3}$) are considered.
\input{bigtable.tex}

\end{document}